The CrypTool Script: \\ Cryptography, Mathematics and More \\ \vspace{30pt} {\large Background reading\\%*[-8pt] for the free e-learning program CrypTool\\ (10th edition -- distributed with CrypTool version 1.4.30)} \vspace{60pt}

### 2013-06-18 script/en/chapters/numbertheory.tex
CT Script: Minor changes including the section about Shared Primes (currently only in English).

\noindent Cycle here means a sequence of numbers $a^i{\rm ~mod~}n$ with $1\leq i<n$
for a given $a$, and a repeating sequence. Within the cycle each number is unique.
Generally, cycles don't have to contain the $1$ -- in the opposite to cycles
which have a multiplicative order $\ge 1$ (they yet have the $1$ always at the end).\\
With $l$ we now mean the cycle length.

@@

% --------------------------------------------------------------------------
\vskip +20pt
\subsubsection{Shared Primes: Modules with common prime factors}
\label{nt_Shared-Primes} \index{Prime!Shared}%\mbox{}
%~\ref{nt_Shared-Primes}, S.~\pageref{nt_Shared-Primes}
In xxxxxxxxxxxxxxxxxxx

@@

\newpage
\hypertarget{nt:AppArith3d}{}
\label{nt:AppArith3d}{}
The following code for generating the tables~\ref{expmod14} and~\ref{expmod22}
at page~\pageref{expmod14} f. also delivers the result in a way, that in can be
easily processed in LaTeX. The prerequisite is that all content is assigned to
one Sage object (here a matrix).\footnote{%
        Remark about the Sage program, especially the Sage
        indices\index{Sage}\index{Sage!latex()}:
        \begin{compactitem}
         \item for x in xrange(2, 5) delivers 2,3,4.
         \item m = matrix(ZZ, 2, 5) has 2 rows and 5 columns.\\
               The cells are named m(0,0) to m(1,4).
         \item All elements of the matrix have to be numerical,
               so ``0'' instead of ``None'' as in the tables before.
         \item The output of matrices kann be controlled in Sage with:
\begin{Verbatim}
               sage: from sage.matrix.matrix0 import set_max_cols, set_max_rows
               sage: set_max_cols(100)
               sage: set_max_rows(100)
\end{Verbatim}
         \item The length of the cycle in the last column of the tables~\ref{expmod14}
               and~\ref{expmod22} was added manually.
        \end{compactitem}
        \vspace{-\baselineskip} % Here its necessary, so that all fit to one page!
        }

@@

\noindent Sage source code~\ref{nt_sagesample_Count_RSA_Keys} below defining the
function \verb#count_Number_of_RSA_Keys# can answer this question concretely
(if the modulus is not too big).\footnote{%
%\begin{compactitem}
\newlength{\saveleftmargini}
\setlength{\saveleftmargini}{\leftmargini}
\setlength{\leftmargini}{0em}% for example to outdent verse
\settowidth{\versewidth}{xxxxx xxxxx xxxxx xxxxx xxxxx xxxxx xxxxx xxxxx xxxxx xxxxx xxxxx xxxxx xxxxx xxxxx xxxxx xxxxx xxxxx xxxxx}
\vspace{-\baselineskip} % TODO: Geht das nicht anders?
\begin{verse}[\versewidth]
 a) Calling \verb#sage: count_Number_of_RSA_Keys(100, 1000)# means to consider the interval
$[100, 1000]$ for $n$.
\verselinebreak $n$ is defined by the two primes $p, q: n = p*q$.
\verselinebreak So here one prime can have the maximal value $500$ because $2 * 500 =1000$
(while then the other prime will have the smallest possible prime value $2$).\\
\vin The number of possible combinations of primes is $comb = 258$.\\
\vin The number of primes in the given range is $143$.\\
\vin The number of private keys is $34,816$.
\end{verse}
%\noindent %       \item
\begin{verse}[\versewidth]
 b) Calling \verb#sage: count_Number_of_RSA_Keys(100, 100, True)# has the
following output:\\
\vin    - Number of private keys for modulus in a given range: 0\\
\vin    - Number of primes in a given range: 0\\
\vin    The reason for that is, that with this call only $n=100$ is considered, and the
   function investigates only\\
\vin semiprime $n$: $100$ is not semi prime\index{Prime number!semi prime}\index{Prime number!half prime}\index{Half prime},
   this means $100$ is not the product of only two primes.
\end{verse}
\setlength{\leftmargini}{\saveleftmargini}% restore original value%\end{compactitem}
}

@@

print "\nRemark: You can verify each fixpoint with power_mod(m,e,n).";
\end{Verbatim}
\caption{Determining all Fixed Points for a Specific Public RSA Key}
\label{nt_sagesample_Calculate_RSA-Fixpoints}
\end{sagecode}
%#print "        Here done for the last found fixpoint:";
%#m = crt(Integer(r),Integer(s),Integer(p),Integer(q))
%#print "        m = ", m, ",", "power_mod = ", power_mod(m,e,n)
%#if (m != power_mod(m,e,n)):
%#                print "Verification failed !!!";

@@

@@

@@

@@

@@

>>>>

\noindent Cycle here means a sequence of numbers $a^i{\rm ~mod~}n$ with $1\leq
i<n$ for a given $a$, and a repeating sequence. According to the generation
method as modular power, here each number is unique within a cycle.
The cycles here don't have to contain the $1$ -- unless this cycles belongs to
a multiplicative order $\ge 1$ (they have the $1$ always at the end of the
cycle and at the position $a^{n-1} ~mod~n$).\\
\noindent With $l$ we now mean the cycle length.

@@

% --------------------------------------------------------------------------
% \vskip +20pt
\newpage
\subsubsection{Shared Primes: Modules with common prime factors}
\label{nt_Shared-Primes} \index{Prime!Shared}%\mbox{}
%~\ref{nt_Shared-Primes}, page~\pageref{nt_Shared-Primes}

The RSA algorithm for public-key cryptography is based on the presumed difficulty of factoring large bi-prime integers, the factoring problem. However, as pointed out in Lenstra et al \cite{nt:Lenstra2012} it is possible, given a set of moduli, to factor some of them by finding shared primes. This way the factoring problem is bypassed using a simple greatest common divisor (gcd) operation. It is no trivial task to extract common shared primes and to factor the according moduli efficiently for a very big number of given moduli (several millions).

Using the gcd only works if the RSA keys were not generated randomly. Taking into consideration the significance of strong cryptographic keys it is important to verify that all keys were generated following the principle of true randomness \cite{nt:Esslinger2012}.

When Lenstra et al published their paper \cite{nt:Lenstra2012} in Feb. 2012, they did not publish the source code. However, soon afterwards the source code of a similar program was published at the CrypTool website\footnote{%
\url{http://www.cryptool.org/en/ctp-dokumentation-en/361-ctp-paper-rsa-moduli}
%\url{http://www.cryptool.org/de/ctp-dokumentation-de/361-ctp-paper-rsa-moduli}
}
in Python and C++, and at the page used by \cite{nt:Heninger2012}\footnote{\url{https://www.factorable.net/}
}.
The fastest code known to me comes with \cite{nt:Heninger2012}.

These applications find all shared factors that may exist, given a finite set of moduli -- even if this set includes millions of moduli. Such an application enables system administrators to test their own RSA keys.

The quite naive way to find all shared factors would be to compare each modul with all other moduli which has a complexity growing quadratically with the number of moduli.
A very efficient method using trees for comparing all gcd pairs is based on a publication of Dan Bernstein in 2005 \cite{nt:Bernstein2005}. Bernstein uses a precalculation which leads to the product of all moduli. It's another example showing how helpful precalculations can be to break cryptographic systems (another famous example are rainbow tables used to find the origin of a hash value \cite{nt:Oechslin2003}).

The following Sage sample shows the very different run times when calculating a gcd and a factorization. The section after this sample will explain the essential part of the method used in \cite{nt:Heninger2012}: Using the trees accelerates the calculation of the gcd pairs a lot.

The Sage sample~\ref{nt_sagesample_Compare-Runtime-gcd-factoring} shows that multiplication of factors, dividing a modul with a known factor, or calculating the gcd is very fast. However, factoring moduli steeply increases with longer moduli. Even the relatively small moduli used in this example show this: The smaller modul (69 decimal digits, 228 bits) took 76 seconds, while the bigger one (72 decimal digits, 239 bits) took almost 217 seconds.

In addition, the operations multiplication, divsion and gcd show big differences in runtime when the used operands are very different in size.

% Just to show the origin of the used primes
% sage: factor (2^211-1)
% 15193 * 60272956433838849161 * 3593875704495823757388199894268773153439
%
% sage: factor (2^214-1)
% 3 * 643 * 84115747449047881488635567801 * 162259276829213363391578010288127
\begin{sagecode}
\begin{Verbatim}%
[fontsize=\footnotesize,fontshape=tt]

# Multiplication
sage: 3593875704495823757388199894268773153439 * 84115747449047881488635567801
302301541122639745170382530168903859625492057067780948293331060817639

sage: 3593875704495823757388199894268773153439 * 162259276829213363391578010288127
583139672825572068433667900695808357466165186436234672858047078770918753

# Division
sage: time 302301541122639745170382530168903859625492057067780948293331060817639 / 
           3593875704495823757388199894268773153439
Wall time: 0.00 s
84115747449047881488635567801

sage: time 583139672825572068433667900695808357466165186436234672858047078770918753 / 
           3593875704495823757388199894268773153439
Wall time: 0.00 s
162259276829213363391578010288127

# Calculate gcd
sage: time gcd (583139672825572068433667900695808357466165186436234672858047078770918753,
                302301541122639745170382530168903859625492057067780948293331060817639)
Wall time: 0.00 s
3593875704495823757388199894268773153439

# Factorize
sage: time factor (583139672825572068433667900695808357466165186436234672858047078770918753)
Wall time: 217.08 s
162259276829213363391578010288127 * 3593875704495823757388199894268773153439

sage: time factor (302301541122639745170382530168903859625492057067780948293331060817639)
Wall time: 76.85 s
84115747449047881488635567801 * 3593875704495823757388199894268773153439

\end{Verbatim}
\caption{Comparing the runtime of calculating a gcd and performing a factorization}
\label{nt_sagesample_Compare-Runtime-gcd-factoring}
\end{sagecode}

\clearpage
\section*{Efficient computing of all gcd pairs and explanation of the formula used}

The paper "Mining Your Ps and Qs: Detection of Widespread Weak Keys in Network Devices"~\cite{nt:Heninger2012} explains the algorithm how the gcd's of every pair of RSA moduli are calculated efficiently.

First the product $P$ of all moduli $m_{i}$ is calculated using a product tree. Then a remainder tree is build modulo the squares of the moduli. Then the gcd's of a defined modul $m_{i}$ and of the remainders $z_{i}$ divided by this defined modul are calculated.

This is visualized in Figure \ref{Figure_Bernstein_Computing-all-pairs-GCDs} which is a copy from \cite{nt:Heninger2012} (where the moduli are called $N_{i}$
instead of $m_{i}$:
\begin{figure}[ht]
\begin{center}
\includegraphics[scale=0.7]{figures/Bernstein_Computing-all-pairs-GCDs.png}
\caption{Algorithm and Figure to compute all gcd pairs efficiently} 
\label{Figure_Bernstein_Computing-all-pairs-GCDs}
\end{center}
\end{figure}

The paper~\cite{nt:Heninger2012} explains well \textit{how} the algorithm works, but not as well \textit{why}. The product $P$ of all moduli is a very big number, even compared to a single modul. Without the simplifications from the remainder tree you would go the following way: Calculate $gcd_{i} = gcd( P / m_{i},  m_{i}) $ for all i. Compare each $gcd_{i} \ne 1$ with all other $gcd_{j} \ne 1$ with  $j \ne i$. If two gcd's are the same, then their moduli share a factor.

As it's very slow to calculate this for numbers with such a big difference in size, the remainder tree is used. Despite it seems to consist of more steps it's a huge simplification.

Within the remainder tree you get -- at the end -- $(P \bmod ({m_{i}^{2}}) ) / m_{i}$ for all $i$.\footnote{%
It would not make sense to calculate $(P \bmod m_{i} ) / m_{i}$
as $(P \bmod m_{i} )$ is always $= 0$.

Sample with very small moduli:\\
$ m_{1} = 2*3 = 6;~~~ m_{2} = 2*7 = 14;~~~ P=6*14=84 $\\
$ P \bmod m_{1} = 84 \bmod 6 = 0;~~~ P \bmod {m_{1}^{2}} = 84 \bmod 36 = 12 $\\
$ P \bmod m_{2} = 84 \bmod 14 = 0;~~~ P \bmod {m_{2}^{2}} = 84 \bmod 196 = 84 $\\
$gcd_{1} = gcd(12/6, 6) = gcd(2, 6) = 2 $\\
$gcd_{2} = gcd(84/14, 14) = gcd(6, 14) = 2 $

The way the tree is structured it also would not make sense to first divide and then do the modulo calculation, as making the division first would lead just to the given moduli but in reversed order.

It also would not make sense to calculate $(P \bmod ({m_{i}^{3}}) ) / {m_{i}^{2}})$ as this is only additional effort with no improvement.
}

The main remaining question now is: Why does $gcd((P \bmod {m_{i}^{2}})/ m_{i}, m_{i})$ deliver the same result as $gcd( P / m_{i}, m_{i})$ ?
We prove that this identity is correct.

P represents here the product of all moduli, and $m_{i}$ represents any
arbitrary modulus.%\mbox{}\\\mbox{}\\

$$gcd((P \bmod {m_{i}^{2}})/ m_{i}, m_{i})~~~  \overset{!}{=}  ~~~gcd( P / m_{i}, m_{i})$$
$ \Longleftrightarrow $~~~\footnote{According to Euklid's algorithm the following identities are true:
\mbox{}\\
$gcd((P \bmod {m_{i}^{2}})/ m_{i}, m_{i})=gcd((P \bmod {m_{i}^{2}})/ m_{i} \bmod{m_{i}}, m_{i})$ \mbox{}\\
$gcd( P / m_{i}, m_{i})=gcd( P / m_{i} \bmod{m_{i}}, m_{i})$\mbox{}
}

$$gcd(((P \bmod {m_{i}^{2}})/ m_{i}) \bmod {m_{i}}, m_{i})~~~  \overset{!}{=}  ~~~gcd( (P / m_{i}) \bmod {m_{i}}, m_{i})$$
$ \Longleftrightarrow $~~~\footnote{The gcd's are equal if both their first arguments are equal.
}

$$((P \bmod {m_{i}^{2}})/ m_{i}) \bmod {m_{i}}~~~  \overset{!}{=}  ~~~ (P / m_{i}) \bmod {m_{i}}$$
$ \Longleftrightarrow $~~~\footnote{The following transformations are all equalities.
}

$$(P \bmod{m_{i}^{2}})/ m_{i} - P / m_{i} \equiv 0   \bmod{m_{i}}~~~ \Leftrightarrow ~~~  m_{i} ~~ | ~~ ((P \bmod{m_{i}^{2}})/ m_{i} - P / m_{i})$$~~~\footnote{%
According to the definition of modulus and division it is:\\
$a \bmod{b} ~~ \Longleftrightarrow ~~ a - b * \lfloor a/b \rfloor$\\
So $P \bmod{m_{i}^{2}}$ can be written as $ P-m_{i}^{2} \lfloor P/m_{i}^{2} \rfloor$.
}

$$ m_{i} ~~ | ~~ ((P-m_{i}^{2}* \lfloor P/m_{i}^{2} \rfloor - P))/m_{i}$$~~~\footnote{%
$P$ reduces itself, the exponent in the $m_{i}$ enumerator is simplified with the  $m_{i}$ denominator.
}

$$ m_{i} ~~ | ~~ (m_{i}* \lfloor P/m_{i}^{2} \rfloor)$$

As this is true, we can conclude that the two gcds are equivalent.

% yyyyyyyyyyyyyyyyyyyyyyyyyyy

% Sätze und Links:
% https://en.wikipedia.org/wiki/Greatest_common_divisor
% https://de.wikipedia.org/wiki/Gr%C3%B6%C3%9Fter_gemeinsamer_Teiler

%TODO: Wohl durch das verschobene Sage-Beispiel wirkt das \newpage vor dem ff. Kapitel nicht!

@@

\newpage
\hypertarget{nt:AppArith3d}{}
\label{nt:AppArith3d}{}
The following code for generating the tables~\ref{expmod14} and~\ref{expmod22}
at page~\pageref{expmod14} f. also delivers the result in a way, that in can be
easily processed in LaTeX. The prerequisite is that all content is assigned to
one Sage object (here a matrix).\footnote{%
        Remark about the Sage program, especially the Sage
        indices\index{Sage}\index{Sage!latex()}:
        \begin{compactitem}
         \item for x in xrange(2, 5) delivers 2,3,4.
         \item m = matrix(ZZ, 2, 5) has 2 rows and 5 columns.\\
               The cells are named m(0,0) to m(1,4).
         \item All elements of the matrix have to be numerical,
               so ``0'' instead of ``None'' as in the tables before.
         \item The output of matrices can be controlled in Sage with:
\begin{Verbatim}
               sage: from sage.matrix.matrix0 import set_max_cols, set_max_rows
               sage: set_max_cols(100)
               sage: set_max_rows(100)
\end{Verbatim}
         \item The length of the cycle in the last column of the tables~\ref{expmod14}
               and~\ref{expmod22} was added manually.
        \end{compactitem}
        \vspace{-\baselineskip} % Here its necessary, so that all fit to one page!
        }

@@

\noindent Sage source code~\ref{nt_sagesample_Count_RSA_Keys} below defining the
function \verb#count_Number_of_RSA_Keys# can answer this question concretely
(if the modulus is not too big).\footnote{%
%\begin{compactitem}
\newlength{\saveleftmargini}
\setlength{\saveleftmargini}{\leftmargini}
\setlength{\leftmargini}{0em}% for example to outdent verse
\settowidth{\versewidth}{xxxxx xxxxx xxxxx xxxxx xxxxx xxxxx xxxxx xxxxx xxxxx xxxxx xxxxx xxxxx xxxxx xxxxx xxxxx xxxxx xxxxx xxxxx}
\vspace{-\baselineskip} % TODO: Geht das nicht anders?
\begin{verse}[\versewidth]
 a) Calling \verb#sage: count_Number_of_RSA_Keys(100, 1000)# means to consider the interval
$[100, 1000]$ for $n$.
\verselinebreak $n$ is defined by the two primes $p, q: n = p*q$.
\verselinebreak So here one prime can have the maximal value $500$ because $2 * 500 =1000$
(while then the other prime will have the smallest possible prime value $2$).\\
\vin The number of possible combinations of primes is $comb = 258$.\\
\vin The number of primes in the given range is $143$.\\
\vin The number of private keys is $34,816$.
\end{verse}
%\noindent %       \item
\begin{verse}[\versewidth]
 b) Calling \verb#sage: count_Number_of_RSA_Keys(100, 100, True)# has the
following output:\\
\vin    - Number of private keys for modulus in a given range: 0\\
\vin    - Number of primes in a given range: 0\\
\vin    The reason for that is, that with this call only $n=100$ is considered,
   and the function investigates only\\
\vin semiprime $n$: $100$ is not semi prime\index{Prime number!semi
   prime}\index{Prime number!half prime},
   this means $100$ is not the product of only two primes.
\end{verse}
\setlength{\leftmargini}{\saveleftmargini}% restore original value%\end{compactitem}
}

@@

print "\nRemark: You can verify each fixpoint with power_mod(m,e,n).";
\end{Verbatim}
\caption{Determining all fixed points for a specific public RSA key}
\label{nt_sagesample_Calculate_RSA-Fixpoints}
\end{sagecode}
%#print "        Here done for the last found fixpoint:";
%#m = crt(Integer(r),Integer(s),Integer(p),Integer(q))
%#print "        m = ", m, ",", "power_mod = ", power_mod(m,e,n)
%#if (m != power_mod(m,e,n)):
%#                print "Verification failed !!!";

@@

\bibitem[Bernstein2005]{nt:Bernstein2005} \index{ Bernstein 2005}
	Daniel J. Bernstein, \\
	{\em Factoring into coprimes in essentially linear time}, \\
	In Journal of Algorithms 54 (2005), 2005,
	\url{http://cr.yp.to/lineartime/dcba-20040404.pdf}.

@@

\bibitem[Esslinger2012]{nt:Esslinger2012} \index{ Esslinger 2012}
	Esslinger, Schneider, and Simon, \\
	{\em RSA -- Sicherheit in der Praxis }, \\
	in KES -- Zeitschrift für Informationssicherheit,
	April 2012.

@@

\bibitem[Heninger2012]{nt:Heninger2012} \index{ Heninger 2012}
	Nadia Heninger, Zakir Durumeric, Eric Wustrow, and J. Alex Halderman, \\
	{\em Mining Your Ps and Qs: Detection of Widespread Weak Keys in
        Network Devices }, \\
	August 2012, \url{https://factorable.net/paper.html}.

@@

\bibitem[Lenstra2012]{nt:Lenstra2012} \index{ Lenstra 2012}
	Arjen K. Lenstra, James P. Hughes, Maxime Augier, Joppe W. Bos,
        Thorsten Kleinjung and Christophe Wachter, \\
	{\em Ron was wrong, Whit is right, A Sanity Check of Public Keys
             Collected on the Web}, \\
	February 2012,
	\url{http://eprint.iacr.org/2012/064.pdf}.

@@

\bibitem[Oechslin2003]{nt:Oechslin2003} \index{ Oechslin 2003}
	Philippe Oechslin,\\
	{\em Making a Faster Cryptanalytic Time-Memory Trade-Off},\\
	At Crypto 2003, 2003,\\
	\url{http://lasecwww.epfl.ch/pub/lasec/doc/Oech03.pdf}.

### 2013-06-18 script/en/chapters/paper_and_pencil.tex
CT Script: Minor changes including the section about Shared Primes (currently only in English).

\item {\bf Rail Fence cipher}\footnote{%
   This method can directly be found in CT1\index{CrypTool 1} at the menu item
   {\bf Encrypt/Decrypt \textbackslash{} Symmetric (classic) \textbackslash{}
   Scytale / Rail Fence}.
   You can simulate this method also under the menu {\bf Encrypt/Decrypt
   \textbackslash{} Symmetric (classic) \textbackslash{} Permutation}:
   For a Rail Fence with 2 lines use as key ``B,A'' and accept the default
   settings (only one permutation, where your input is done line-by-line and
   the output is taken column-by-column). 
   Using the key ``A,B'' would start the zigzag pattern below in the
   way, that the first letter is written into the first line instead of the
   second line.}
   \cite{pp:Singh1999}\index{Rail Fence cipher}:
   The characters of a message are alternately written in two (or more) lines,
   creating a zigzag pattern. The resulting ciphertext is read out 
   line by line.\\
   This is more a children's method.

@@

%\newpage % Damit die Tables immer direkt nach dem zugehörigen Item. 
\item {\bf Cadenus cipher} \cite{pp:ACA2002}\index{Cadenus}: Cadenus is a form
   of columnar transposition that uses two keywords.\\
   The 1st keyword is used to swap columns.\\
   The 2nd keyword is used to define the initial letter of each column:
   this 2nd keyword is a permutation of the used alphabet. This permutation
   is written on the left of the first column.
   Afterwards, each column is moved (wrap-around) so that it begins with the
   letter, which is in the same line as the key letter of the first 
   keyword within the second keyword. Ciphertext is read out line by line.

@@

\item {\bf Atbash cipher\footnotemark}
   \footnotetext{%
   This cipher can be simulated in CT1\index{CrypTool 1} 
   under the menu {\bf Encrypt/Decrypt \textbackslash{} Symmetric (classic) 
   \textbackslash{} Substitution / Atbash}.}  
   \cite{pp:Singh1999}\index{Atbash}: Replace the first letter of the 
   alphabet by the last letter of the alphabet, the second one by the last 
   but one, etc.

\item {\bf Shift cipher, for example Caesar cipher}\footnote{In CT1\index{CrypTool 1}
   this method  can be found at three different places in the menu tree:\\
   - {\bf Encrypt/Decrypt \textbackslash{} Symmetric (classic) \textbackslash{} 
      Caesar / ROT13}\\
   - {\bf Analysis \textbackslash{} Symmetric Encryption (classic)
     \textbackslash{} Ciphertext only \textbackslash{} Caesar} \\
   - {\bf Indiv. Procedures \textbackslash{} Visualization of Algorithms 
     \textbackslash{} Caesar}. } 
   \cite{pp:Singh1999}\index{Caesar}%
   : Plaintext alphabet and ciphertext alphabet are shifted against each 
   other by a determined number of letters. Using the Caesar cipher means 
   shifting letters about three positions.

@@

\item {\bf Straddling Checkerboard} 
   \cite{pp:Goebel2003}\index{Straddling Checkerboard}:
   A 3x10 matrix is filled
   with the letters of the used alphabet and two arbitrary digits or special
   characters as follows: The different letters of a keyword and the remaining
   characters are written into the grille. The columns are numbered 0 to 9, the
   second and the third line are numbered 1 and 2. Each plaintext character is
   replaced by the corresponding digit, respectively the corresponding pair of
   digits. As ``1'' and ``2'' are the first digits of the possible 
   two-digit-numbers, they are not used as single digits.

@@

\item {\bf Straddling Checkerboard, variant} \cite{pp:Goebel2003}
   \index{Straddling Checkerboard}: This variant
   of the straddling checkerboard was developed by Soviet spies during WW2. 
   Ernesto (Ch\'e) Guevara \index{Ch\'e Guevara} and Fidel Castro allegedly
   used this cipher for their secret communication.
   A grille is filled with the alphabet (number of columns = length of
   keyword), and two arbitrary digits are chosen as reserved to indicate the
   second and third line of a 3x10-matrix (see above). Now the grille is
   traversed column by column and the single letters are transferred row by
   row into the matrix: 
   For a faster encryption, the eight most common letters (ENIRSATO) are 
   assigned the digits from 0 to 9, the reserved 2 digits are not assigned. 
   The remaining letters are provided with combinations of digits one after
   another and are inserted into the grille.

@@

%------------------------------------------------------------------------------
\section{Combining substitution and transposition}
\label{PaP_combined_ciphers}
\index{Product algorithm}\index{Encryption!Product algorithm}%
\index{Cascade cipher}\index{Encryption!Cascade cipher}%

@@

% ---------------------------------------------------------------------------
\newpage
\subsubsection{Caesar cipher}
\index{Caesar}

>>>>

\item {\bf Rail fence cipher}\footnote{%
   This method can directly be found in CT1\index{CrypTool 1} at the menu item
   {\bf Encrypt/Decrypt \textbackslash{} Symmetric (classic) \textbackslash{}
   Scytale / Rail Fence}.
   You can simulate this method also under the menu {\bf Encrypt/Decrypt
   \textbackslash{} Symmetric (classic) \textbackslash{} Permutation}:
   For a Rail Fence with 2 lines use as key ``B,A'' and accept the default
   settings (only one permutation, where your input is done line-by-line and
   the output is taken column-by-column). 
   Using the key ``A,B'' would start the zigzag pattern below in the
   way, that the first letter is written into the first line instead of the
   second line.}
   \cite{pp:Singh1999}\index{Rail fence cipher}:
   The characters of a message are alternately written in two (or more) lines,
   creating a zigzag pattern. The resulting ciphertext is read out 
   line by line.\\
   This is more a children's method.

@@

%\newpage % Damit die Tables immer direkt nach dem zugehörigen Item. 
\item {\bf Cadenus cipher} \cite{pp:ACA2002}\index{Cadenus cipher}: Cadenus is
   a form of columnar transposition that uses two keywords.\\
   The 1st keyword is used to swap columns.\\
   The 2nd keyword is used to define the initial letter of each column:
   this 2nd keyword is a permutation of the used alphabet. This permutation
   is written on the left of the first column.
   Afterwards, each column is moved (wrap-around) so that it begins with the
   letter, which is in the same line as the key letter of the first 
   keyword within the second keyword. Ciphertext is read out line by line.

@@

\item {\bf Atbash cipher\footnotemark}
   \footnotetext{%
   This cipher can be simulated in CT1\index{CrypTool 1} 
   under the menu {\bf Encrypt/Decrypt \textbackslash{} Symmetric (classic) 
   \textbackslash{} Substitution / Atbash}.}  
   \cite{pp:Singh1999}\index{Atbash cipher}: Replace the first letter of the 
   alphabet by the last letter of the alphabet, the second one by the last 
   but one, etc.

\item {\bf Shift cipher, for example Caesar cipher}\footnote{In CT1\index{CrypTool 1}
   this method  can be found at three different places in the menu tree:\\
   - {\bf Encrypt/Decrypt \textbackslash{} Symmetric (classic) \textbackslash{} 
      Caesar / ROT13}\\
   - {\bf Analysis \textbackslash{} Symmetric Encryption (classic)
     \textbackslash{} Ciphertext only \textbackslash{} Caesar} \\
   - {\bf Indiv. Procedures \textbackslash{} Visualization of Algorithms 
     \textbackslash{} Caesar}. } 
   \cite{pp:Singh1999}\index{Caesar cipher}%
   : Plaintext alphabet and ciphertext alphabet are shifted against each 
   other by a determined number of letters. Using the Caesar cipher means 
   shifting letters about three positions.

@@

\item {\bf Straddling Checkerboard} 
   \cite{pp:Goebel2003}\index{Straddling checkerboard}:
   A 3x10 matrix is filled
   with the letters of the used alphabet and two arbitrary digits or special
   characters as follows: The different letters of a keyword and the remaining
   characters are written into the grille. The columns are numbered 0 to 9, the
   second and the third line are numbered 1 and 2. Each plaintext character is
   replaced by the corresponding digit, respectively the corresponding pair of
   digits. As ``1'' and ``2'' are the first digits of the possible 
   two-digit-numbers, they are not used as single digits.

@@

\item {\bf Straddling Checkerboard, variant} \cite{pp:Goebel2003}
   \index{Straddling checkerboard}: This variant
   of the straddling checkerboard was developed by Soviet spies during WW2. 
   Ernesto (Ch\'e) Guevara \index{Ch\'e Guevara} and Fidel Castro allegedly
   used this cipher for their secret communication.
   A grille is filled with the alphabet (number of columns = length of
   keyword), and two arbitrary digits are chosen as reserved to indicate the
   second and third line of a 3x10-matrix (see above). Now the grille is
   traversed column by column and the single letters are transferred row by
   row into the matrix: 
   For a faster encryption, the eight most common letters (ENIRSATO) are 
   assigned the digits from 0 to 9, the reserved 2 digits are not assigned. 
   The remaining letters are provided with combinations of digits one after
   another and are inserted into the grille.

@@

%------------------------------------------------------------------------------
\section{Combining substitution and transposition}
\label{PaP_combined_ciphers}
\index{Product algorithm}\index{Encryption!product algorithm}%
\index{Cascade cipher}\index{Encryption!cascade cipher}%

@@

% ---------------------------------------------------------------------------
\newpage
\subsubsection{Caesar cipher}
\index{Caesar cipher}

### 2013-06-18 script/en/chapters/primes.tex
CT Script: Minor changes including the section about Shared Primes (currently only in English).

     Green and Tao intended to proof that there are infinitely many arithmetic
     sequences of length four. For this they considered sets of numbers 
     consisting of primes and so called 
     ``near primes''\index{Near prime}\index{Prime number!near prime}.
     These are numbers with a small set of divisors like numbers which are
     the product of two primes - these numbers are called 
     ``half primes''%
     \index{Half prime}\index{Prime number!half prime}%
     \index{Prime number!semi prime}\index{Prime number!bi prime}%
     \index{Number!semi prime}\index{Number!bi prime}.
     Thus they managed to considerably simplify their work because about near
     primes there already existed a lot of useful theorems.  Finally they
     discovered that the results of their theorem were far more reaching than
     they had assumed and so they were able to prove Hardy's conjecture.

@@

     If someone wants to discover such sequences he should consider the
     following thought.
     The length of a sequence determines the minimal common distance
     between the single primes of the sequence. Given a sequence with 6 
     elements the distance between them has to be 30 or a multiple of 30.
     The number 30 results as the product of all primes smaller than the
     length of the sequence. So its the product of all primes smaller than 6:
     $ 6\# = 5\# = 2 * 3 * 5 = 30 $.
     It is $ 10\# = 7\# = 2 ∑ 3 ∑ 5 ∑ 7 = 210$.
     If you look for a sequence with 15 elements, then the common
     distance is at least $15\# = 13\# = 2 * 3 * 5 * 7 * 11 * 13 = 30,030$.
     \index{Primzahl!k Primorial} \index{Primzahl!k\#}

@@

Here we will have a closer look at what is called the\index{Goldbach Conjecture}
     \textbf{Goldbach Conjecture}\footnote{%
        \url{http://en.wikipedia.org/wiki/Goldbach%27s_conjecture} }.

@@

\vskip +25 pt
\hypertarget{HT-GoogleRecruitment2004}{}
\subsubsection{Recruitment at Google in 2004}%
% \paragraph*{Recruitment at Google in 2004}%
\index{Google!Recruitment}%
\label{HT-GoogleRecruitment2004}%
\mbox{}
In summer 2004 the company Google used the number $e$ to attract potential
employees.\footnote{The base of the natural logarithm\index{logarithm!natural}
$e$ is approximately 2.718 281 828 459.
This is one of the most important numbers in all of mathematics like complex
analysis, finance, physics and geometry. Now it was used the first time -- as
far as I know -- for marketing or recruitment.}${}^,$\footnote{Most of this
information is taken from the article ``e-number crunching'' by John Allen
Paulos in TheGuardian, Sept. 30, 2004, and from the web:\\
- \url{http://www.mkaz.com/math/google/}\\
- \url{http://epramono.blogspot.com/2004/10/7427466391.html}\\
- \url{http://mathworld.wolfram.com/news/2004-10-13/google/}\\
- \url{http://www.math.temple.edu/~paulos/}.
}

>>>>

     Green and Tao intended to proof that there are infinitely many arithmetic
     sequences of length four. For this they considered sets of numbers 
     consisting of primes and so called 
     ``near primes''\index{Prime number!near prime}.
     These are numbers with a small set of divisors like numbers which are
     the product of two primes - these numbers are called 
     ``half primes''%
     \index{Prime number!half prime}%
     \index{Prime number!semi prime}\index{Prime number!bi prime}%
     \index{Number!semi prime}\index{Number!bi prime}.
     Thus they managed to considerably simplify their work because about near
     primes there already existed a lot of useful theorems.  Finally they
     discovered that the results of their theorem were far more reaching than
     they had assumed and so they were able to prove Hardy's conjecture.

@@

     If someone wants to discover such sequences he should consider the
     following thought.
     The length of a sequence determines the minimal common distance
     between the single primes of the sequence. Given a sequence with 6 
     elements the distance between them has to be 30 or a multiple of 30.
     The number 30 results as the product of all primes smaller than the
     length of the sequence. So its the product of all primes smaller than 6:
     $ 6\# = 5\# = 2 * 3 * 5 = 30 $.
     It is $ 10\# = 7\# = 2 ∑ 3 ∑ 5 ∑ 7 = 210$.
     If you look for a sequence with 15 elements, then the common
     distance is at least $15\# = 13\# = 2 * 3 * 5 * 7 * 11 * 13 = 30,030$.
     \index{Prime number!k primorial} \index{Prime number!k\#}

@@

Here we will have a closer look at what is called the\index{Goldbach conjecture}
     \textbf{Goldbach conjecture}\footnote{%
        \url{http://en.wikipedia.org/wiki/Goldbach%27s_conjecture} }.

@@

\vskip +25 pt
\hypertarget{HT-GoogleRecruitment2004}{}
\subsubsection{Recruitment at Google in 2004}%
% \paragraph*{Recruitment at Google in 2004}%
\index{Google!Recruitment}%
\label{HT-GoogleRecruitment2004}%
\mbox{}
In summer 2004 the company Google used the number $e$ to attract potential
employees.\footnote{The base of the natural logarithm\index{Logarithm!natural}
$e$ is approximately 2.718 281 828 459.
This is one of the most important numbers in all of mathematics like complex
analysis, finance, physics and geometry. Now it was used the first time -- as
far as I know -- for marketing or recruitment.}${}^,$\footnote{Most of this
information is taken from the article ``e-number crunching'' by John Allen
Paulos in TheGuardian, Sept. 30, 2004, and from the web:\\
- \url{http://www.mkaz.com/math/google/}\\
- \url{http://epramono.blogspot.com/2004/10/7427466391.html}\\
- \url{http://mathworld.wolfram.com/news/2004-10-13/google/}\\
- \url{http://www.math.temple.edu/~paulos/}.
}
